Issues with Dynamical Measurements Across the Disk

### 2019-02-12 paper/ms.tex
edits to the second half of 2.3, some suggestions for terminology changes, and a new proposal for the name of section 3

\begin{deluxetable*}{ccccccc}
\tablecaption{Description and names of the three orbits considered in this
work.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$}}
\startdata 
thin & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
thick & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
\enddata
\end{deluxetable*}

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the offset coordinate systems depend on the time (i.e. orbital phase) at which the star's phase-space position is observed. (To our extremely long-lived observer this would appear as time-dependence in the actions, which appear to oscillate around their value using the correct coordinate system even though in this example the Galaxy is described perfectly by a static, axisymmetric potential.) The relative size of the phase variation in each action depends on the direction of the systematic offset as well as the true values of the actions (i.e. the type of orbit). For a systematic offset in $z$ ({\em upper} panels), the 95\textsuperscript{th} minus
5\textsuperscript{th} percentiles are $2.2$ and $6.2 \actunit$ for $J_r$ and
$J_z$, respectively. These are fractional errors of $5.7\%$ and $85.7\%$,
respectively. The error induced in $L_z$ is negligible, as expected since
$L_z$ is independent of $z$. It is worth pointing out that a
$100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced a $43\%$ error in the computation of $J_z$.

@@

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick disk
orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $L_z$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_r$, with the middle
$90\%$ of values over the $\Gyr$ being $\sim8\%$ of the true $J_r$. As
expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
with the middle $90\%$ of values being $\sim43\%$ of the true $J_z$. The $x$
offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

We now repeat the procedure but continuously changing the error in the $z$ and
$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
minus 5\uth percentile divided by the true action value for three different
orbits. The thick (\thickcolor) orbit is exactly the same as in
Figure~\ref{fig:one_orbit_wrong_ref}, and we also consider the thin and halo
orbits (\thincolor and \halocolor, respectively) from Table~\ref{tab:orbits}.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the error in each
action assuming an offset in the $z$ component (i.e. the Galactic midplane).
In the {\em bottom row} we also consider offsets in the $x$ component (i.e.
the Solar radius). The {\em left}, {\em middle}, and {\em right} columns show
the relative error in $J_z$, $L_z$, and $J_r$, respectively. We compute the
relative error as the  95\uth minus 5\uth percentile of the measured action
($\Delta J_i$) over the course of the first $\Gyr$ of the incorrectly observed
orbit divided by the true action value as a percentage.

In the {\em upper middle} panel, there is essentially no error in the
determination of $L_z$. This is expected since, in principle, $L_z$ only
depends on the $x$- and $y$-components of the position and velocity of the
stars\footnote{Though this is not how we compute $L_z$ in practice.}, and is
thus unaffected by offsets in $z$. This also agrees with the result we found
in Figure~\ref{fig:one_orbit_wrong_ref}.

The {\em upper right} panel shows that the fractional error in $J_z$ is more
exaggerated for more planar orbits. For the thin orbit, deviations that are
$\sim30\,\pc$ already give $100\%$ deviations in the actions. The required
offset for $100\%$ deviation is $\sim225\,\pc$ for the thick orbit. The halo
orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
error in $J_z$ out to a midplane offset of $500\,\pc$.

For the offset in the Solar radius {\em bottom panels}, the error is largest
for $J_r$, with some deviations resulting in $L_z$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines overlap.

@@

\section{Local Midplane in FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological zoom-ins used in this work.
Cosmological zoom-ins allow one to simulate a selected region at high
resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. The FIRE
collaboration has simulated a number of Milky Way-mass zoom-ins as part of the
{\em Latte} suite of FIRE-2 simulations
\citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
considered in \citet{2018arXiv180610564S} --- m12i, m12f, m12m, since they
have associated mock {\em Gaia} DR2 catalogues and the $z=0$ snapshots are
publicly available.

>>>>

\begin{deluxetable*}{ccccccc}
\tablecaption{Description and names of the three orbits considered in this
work.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
\enddata
\end{deluxetable*}

@@

Figure~\ref{fig:one_orbit_wrong_ref} shows that the actions computed in the offset coordinate systems depend on the time (i.e. orbital phase) at which the star's phase-space position is observed. (To our extremely long-lived observer this would appear as time-dependence in the actions, which appear to oscillate around their value using the correct coordinate system even though in this example the Galaxy is described perfectly by a static, axisymmetric potential.) The relative size of the phase variation in each action depends on the direction of the systematic offset as well as the true values of the actions (i.e. the type of orbit). In reality we will have one measurement of the phase-space position to work with, in which case the determination of the orbital phase in $z$ or $R$ is degenerate with the degree of systematic offset in that coordinate (see Figure \ref{fig:cartoon}). Additionally, in searching for migrated members of known open clusters, we would compute actions for stars with a wide range of orbital phases and locations in the Galaxy. In the following we therefore quote percentile ranges for the possible values computed for each action as a proxy for the effect of these systematic errors in the coordinate system.

For a systematic offset in $z$ ({\em upper} panels), the 95\textsuperscript{th} minus
5\textsuperscript{th} percentiles are $2.2$ and $6.2 \actunit$ for $J_r$ and
$J_z$, respectively. These are fractional errors of $5.7\%$ and $85.7\%$,
respectively. The error induced in $L_z$ is negligible, as expected since
$L_z$ is independent of $z$. It is worth pointing out that a
$100\,\pc$ error in an orbit with $z_{\text{max}}=850\,\pc$ --- a $12\%$ error
--- induced a $43\%$ error in the computation of $J_z$.

@@

\begin{figure*}
\plotone{fig/schmactions_one_orbit.pdf}
\caption{The artificial phase-dependence in the observed actions induced by an
error in the Galactocentric coordinate system. We consider here the thick disk
orbit, which has actions of $(J_r, L_z, J_z) = (37.9, -1520, 7.0)\,\actunit$
and $z_{\text{max}}=850\,\pc$ (see Table~\ref{tab:orbits}). We integrate the
orbit according to the procedure laid out in Section~\ref{ssec:action_comp}.
Then, we subtract $100\,\pc$ from the $z$ value ({\em upper panels}) or the
$x$ value ({\em lower panels}) of each position in the orbit, corresponding to
an erroneous observer assuming a midplane ({\em upper}) or solar radius ({\em
lower}) that is off by $100\,\pc$. We then allow our (immortal) observer to
observe the orbit over $1\,\Gyr$ and perform the same orbit integration
procedure at each timestep, and report the values of the actions. The
computation of $L_z$ is pristine to errors in $z$, with only numerical
artifacts remaining. Only small errors are induced in $J_r$, with the middle
$90\%$ of values over the $\Gyr$ being within $\sim8\%$ of the true $J_r$. As
expected, large errors are induced in $J_z$ with a $100\,\pc$ offset in $z$,
with the middle $90\%$ of values being within $\sim43\%$ of the true $J_z$. The $x$
offset induces uncertainties in $J_r$ and $L_z$ of $\sim21\%$ and $\sim3\%$,
respectively. A $\sim3\%$ error in $J_z$ is also induced.}
\label{fig:one_orbit_wrong_ref}
\end{figure*}

We now repeat the same procedure for systematic offsets between 0 and 500 pc in the $z$ and
$x$ components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth
minus 5\uth percentile divided by the true action value for the three different fiducial
orbits in Table \ref{tab:orbits}. The thick-disk (\thickcolor) orbit is the one from
Figure~\ref{fig:one_orbit_wrong_ref}, but we also now consider the effect on the action determined for the thin-disk (\thincolor) and halo (\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced in each
action for an offset in the $z$ component (i.e. the Galactic midplane).
In the {\em bottom row} we consider offsets in the $x$ component (i.e.
the Solar radius). The {\em left}, {\em middle}, and {\em right} columns show
the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$, respectively. We compute the
fractional spread $\Delta J_i/J_i$ as the 95\uth minus 5\uth percentile of the action computed using the 
offset coordinate system over the course of the first $\Gyr$, divided by the true action (computed in the correct coordinate system).

In the {\em upper middle} panel, there is essentially no spread in the
determination of $L_z$. This is expected since, in principle, $L_z$ only
depends on the $x$- and $y$-components of the position and velocity of the
stars\footnote{In practice, however, $L_z$ is computed as part of the torus-fitting method.}, and is
thus unaffected by offsets in $z$. Indeed, the result we found
in Figure~\ref{fig:one_orbit_wrong_ref} for the thick-disk orbit holds for all orbit types.

The {\em upper right} panel shows that the fractional error in $J_z$ is more
exaggerated for more planar (disk-like) orbits. For the thin-disk orbit, a systematic offset of
$\sim30\,\pc$ in the $z$ coordinate already results in $100\%$ deviations in the actions. The required
offset for $100\%$ deviation is $\sim225\,\pc$ for the thick-disk orbit. The halo
orbit is relatively resistant to errors in the midplane, with only $\sim25\%$
error in $J_z$ out to a midplane offset of $500\,\pc$.

For the offset in the Solar radius {\em bottom panels}, the error is largest
for $J_r$, with some deviations resulting in $L_z$ and relatively small
deviations in $J_z$. In the {\em bottom middle} and {\em bottom right} panels
all three lines nearly overlap.

@@

%where is Figure 4 referred to? we should include a discussion of it before starting

\section{Azimuthal midplane variations at the Solar Circle: estimates from FIRE} \label{sec:local_fire}
\subsection{Description of Simulation} \label{ssec:cosmozoom}
Here we briefly describe the cosmological zoom-ins used in this work.
Cosmological zoom-ins allow one to simulate a selected region at high
resolution embedded in a low-resolution cosmological background
\citep[e.g.][]{1993ApJ...412..455K,2014MNRAS.437.1894O}. The FIRE
collaboration has simulated a number of Milky Way-mass zoom-ins as part of the
{\em Latte} suite of FIRE-2 simulations
\citep{2016ApJ...827L..23W,2018MNRAS.481.4133G}. We use the three zoom-ins
considered in \citet{2018arXiv180610564S} --- m12i, m12f, m12m, since they
have associated mock {\em Gaia} DR2 catalogues and the $z=0$ snapshots are
publicly available.
